---
title: "Data Wrangling Exercise 1"
author: "Kevin Baca"
date: "3/9/2019"
output: 
  html_document: 
    theme: simplex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project: Basic Data Manipulation
Data wrangling exercise with a toy electronic store dataset. 

####Task Overview
1. [Load the data in RStudio](#task1)
2. [Clean up brand names](#task2)
3. [Separate product code and number](#task3)
4. [Add product categories](#task4)
5. [Add full address for geocoding](#task5)
6. [Create dummy variables for company and product category](#task6)
7. Submit the project on Github


##Task 1 | Load the Data into RStudio {#task1}
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(readr)
refinedatasheet1 <- read_csv("~/Downloads/refinedatasheet1.csv")
refine <- refinedatasheet1
glimpse(refine)
```


##Task 2 | Cleanup Brand Names {#task2}
*Clean up the 'company' column so all of the misspellings of the brand names are standardized. 
*For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).
```{r}
#first I want to get a look at the unique observations in the company columnn
refine %>%  group_by(company) %>% tally

#mutating all of the observations in the company variable to lowercase
refine <- refine %>% 
  mutate(company = tolower(company))

#from tobias: I should the `case_when` construction in tandem with the str_detect function from tidyverse.
#Using this construction, the tilda operator `~` 
#Note: I'm somewhat dubious about using the company variable as sort of else statement
refine <- refine %>%
  mutate(company = case_when(
    str_detect(company,"ph") ~ "philips",
    str_detect(company,"fil") ~ "philips",
    str_detect(company,"ak") ~ "akzo",
    str_detect(company,"outen") ~ "van houten",
    str_detect(company,"ver") ~ "unliver",
    T ~ company))
         
#checking the results
refine %>%  group_by(company) %>% tally
```


##Task 3 | Separate product code and number {#task3}
*Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```{r}
refine %>% select(`Product code / number`)

# I found this on stackoverflow 
#I creared a new function called "PEWP" which uses the strsplit function to split each element in the product code column character vector into two distinct substrings
#note: I could use regx in place for the split, but then no FIXED argument
PEWP <- strsplit(as.character(refine$`Product code / number`), "-", fixed = TRUE)

#creating two new vectors with sapply from the Split elements
prod_code <- sapply(PEWP, "[", 1)
number <- sapply(PEWP, "[", 2)

#attaching the two newly-created vectors to the dataframe to form new columns/variables
refine <- cbind(refine,prod_code,number)

#viewing new dataframe to spotcheck
tbl_df(refine)
```

##Task 4 | Add product categories {#task4}
*You learn that the product codes actually represent the following product categories:
**p = Smartphone
**v = TV
**x = Laptop
**q = Tablet
*In order to make the data more readable, add a column with the product category for each record.

```{r}
#Using mutate if to coerce all variables to chars
#note: It would be good to do this 
refine <- refine %>% 
  mutate_if(is.factor, as.character) %>% 
  glimpse()

#using the str_detect function again, I am mutating data to a new column I named "caregory"
refine <- refine %>%
  mutate(category = case_when(
     str_detect(prod_code,"p") ~ "Smartphone",
     str_detect(prod_code,"v") ~ "TV",
     str_detect(prod_code, "x") ~ "Laptop",
     str_detect(prod_code, "q") ~ "Tablet",
    T ~ prod_code))

# Using Select to view the relevant columns for a spot check
refine %>% select(company,`Product code / number`,prod_code:category)
```


##Task 5 | Add full address for geocoding {#task5}
*You'd like to view the customer information on a map. 
*In order to do that, the addresses need to be in a form that can be easily geocoded. 
*Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.
```{r}
#viewing columns to understand
refine %>% select(address:country)

#using tidyverse function unite to assign concatenated values to a new column names "geo""
refine <- unite(refine,geo,address,city,country,sep = ", ", remove = TRUE)

# Taking a look
refine %>% select(geo)

```

##task 6 | Create dummy variables for company and product category {#task6}

*Both the _company_ name and _product_ category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.,

*Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.

*Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.

```{r}
#to coerce particular columns to factors, I used a stackoverflow tip that required the magrittr library
library(magrittr)
# assigned the columns to cols
cols <- c("company","category")
# used mutate_each function to convert the cols solumns to factors
refine %>%
       mutate_each(funs(factor(.)),cols)

#creates dummy variables and inner join (I don't see the point of a left join here) the spread columns witht the refine table. 
refine %>% 
  mutate(var = 1) %>% 
  spread(company, var, fill = 0, sep = "_") %>% 
  inner_join(refine) %>% 
  select(company,7:10)

#creates dummy variables and inner join (I don't see the point of a left join here) the spread columns witht the refine table. 
refine %>% 
  mutate(var = 1) %>% 
  spread(category, var, fill = 0, sep = "_") %>% 
  inner_join(refine) %>% 
  select(category,7:10)

```

